---
title: "Cross-linguistic Ability Check"
author: "George"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(mirt)
require(tidyverse)
require(ggpubr)

```

We want to compare the IRT-estimated ability vs. age plots for each language model to make sure that nothing is funky with the fits.
(Particularly since Danish and Norwegian comprehension item parameters have shifted distributions compared to the other languages.)

```{r comprehension, echo=F, eval=F}
load("data/multiling_2pl_WG_comp_fits.Rdata")
languages = setdiff(names(coefs), c("Kigiriama", "French (French)"))
all_demo <- tibble()
for(language in languages) {
  load(paste("data/",language,"_WG_data.Rdata", sep=''))
  # sometimes a mismatch between # of subjects in d_demo and d_comp !
  # e.g., English d_comp = 2454 rows, d_demo = 2435 rows (does d_comp have a few longitudinal?)
  # length(intersect(rownames(d_comp), d_demo$data_id))
  d_comp <- d_comp[which(rowSums(d_comp, na.rm=T)!=0 & 
                           rowSums(d_comp, na.rm=T)!=nrow(d_comp)),]
  thetas <- tibble(data_id = as.numeric(rownames(d_comp)), 
                   comp_ability = fscores(models[[language]], method="MAP", 
                          response.pattern = d_comp,
                          append_response.pattern = F)[,c("G")])                
  all_demo <- bind_rows(all_demo, d_demo %>% left_join(thetas))
  if(!models[[language]]@OptimInfo$converged) print(paste(language,"not converged!"))
}
# problem with French (French): maybe because d_comp columns are item_id ?
# or because of a few near-ceiling/-floor participants?
```


```{r production, echo=F, eval=F}
load("data/multiling_2pl_WG_prod_fits.Rdata")
prod_demo <- tibble()
for(language in languages) {
  load(paste("data/",language,"_WG_data.Rdata", sep=''))
  d_prod <- d_prod[which(rowSums(d_prod, na.rm=T)!=0),]
  thetas <- tibble(data_id = as.numeric(rownames(d_prod)), 
                   prod_ability = fscores(models[[language]], method="MAP", 
                          response.pattern = d_prod,
                          append_response.pattern = F)[,c("G")])                
  prod_demo <- bind_rows(prod_demo, d_demo %>% left_join(thetas))
  if(!models[[language]]@OptimInfo$converged) print(paste(language,"not converged!"))
}
# French (French) not converged!
# French (Quebecois) not converged!

all_demo <- all_demo %>% left_join(prod_demo)
save(all_demo, file="data/xling_WG_abilities.Rdata")
```

## Overall Ability Distributions

```{r, echo=F, warning=F, fig.width=8, fig.height=4.5}
load("data/xling_WG_abilities.Rdata")

p1 <- all_demo %>% ggplot( aes(x=comp_ability, fill=language)) +
    geom_density(alpha=0.6, position = 'identity') +
    #geom_vline(aes(xintercept=mean(comp_ability))) + 
    theme_classic() + xlab("Comprehension Ability")

p2 <- all_demo %>% ggplot( aes(x=prod_ability, fill=language)) +
    geom_density(alpha=0.6, position = 'identity') +
    theme_classic() + xlab("Production Ability")

ggarrange(p1, p2, nrow=1, common.legend=T)
```

The ability distributions per language greatly overlap.

## Comprehension Ability vs. Age

Correlation of comprehension ability and age: `r apa::cor_apa(cor.test(all_demo$comp_ability, all_demo$age), print=F)`

```{r, echo=F, fig.width=10, fig.height=10, warning=F}
all_demo %>% ggplot( aes(x=jitter(age), y=comp_ability)) +
  geom_point(alpha=0.6) +
  facet_wrap(. ~ language) + 
  theme_classic() + ylab("Comprehension Ability") +
  geom_smooth(method=lm)
```

## Production Ability vs. Age

Correlation of production ability and age: `r apa::cor_apa(cor.test(all_demo$prod_ability, all_demo$age), print=F)`

```{r, echo=F, fig.width=10, fig.height=10, warning=F}
all_demo %>% ggplot( aes(x=age, y=prod_ability)) +
  geom_point(alpha=0.6) +
  facet_wrap(. ~ language) + 
  theme_classic() + ylab("Production Ability") +
  geom_smooth(method=lm)
```


## Production vs. Comprehension Ability

Correlation of comprehension ability and production ability: `r apa::cor_apa(cor.test(all_demo$comp_ability, all_demo$prod_ability), print=F)`.

```{r, echo=F, fig.width=10, fig.height=10, warning=F}
all_demo %>% ggplot(aes(x=prod_ability, y=comp_ability)) + # color=language
  geom_point(alpha=.3) + 
  facet_wrap(. ~ language) + 
  xlab("Production Ability") + 
  ylab("Comprehension Ability") +
  theme_classic()
```


