---
title: "Appendix"
author: "George Kachergis"
date: "2023-01-25"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=F, message=F)

library(png)
library(grid)
#library(xtable)
require(mirt)
require(tidyverse)
require(ggpubr)
require(tidyboot)
require(here)
library(RColorBrewer)
library(gplots)
library(kableExtra)

html_table_width <- function(kable_output, width){
  width_html <- paste0(paste0('<col width="', width, '">'), collapse = "\n")
  sub("<table>", paste0("<table>\n", width_html), kable_output)
} 

source(here("cross-ling-comparison-helpers.R"))

# long data-frame of all WS production parameters
load(here("data/xling-WSprod-IRTparms.Rdata"))

low_data_langs = c("Persian","Finnish","Kiswahili","Irish",
                   "English (Irish)","Kigiriama","Spanish (Peruvian)","Greek (Cypriot)")

xldf <- xldf %>% mutate(uni_lemma = ifelse(uni_lemma=="NA", NA, uni_lemma))
xldf <- xldf %>% filter(!is.na(uni_lemma)) # only want the words with defined uni-lemmas
xldf_lowd <- xldf %>% filter(is.element(language, low_data_langs)) # 5464
xldf <- xldf %>% filter(!is.element(language, low_data_langs)) # 17715

# align some categories: Russian has descriptive adjectives and adverbs
xldf[which(xldf$category=="descriptive_words (adjectives)"),]$category = "descriptive_words"
xldf[which(xldf$category=="outside_places"),]$category = "outside"


olap <- get_uni_lemma_overlap(xldf)
```

## Uni-lemma Overlap on CDI:WS Forms

Number of overlapping items on each pair of CDI:WS forms.

```{r, fig.width=10, fig.height=10}
# show heatmap of overlap between each CDI:WS
heatmap.2(olap, dendrogram = "none", margins=c(10,10))
```


## Word Difficulty by Semantic Category and Language

```{r semantic-cats-by-lang, fig.width=10, fig.height=18, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Mean difficulty of CDI words by semantic category and language. Bars represent bootstrapped 95\\% confidence intervals."}
prod_cat <- xldf %>% group_by(language, category) %>%
  filter(!is.na(d), !is.na(category),
         !is.element(category, c("final_particles", "directions", "numbers", "articles", "other",
                                 "descriptive_words (adverbs)", "states", "unknown",
                                 "verb_endings", "verb_modifiers",
                                 "classifiers", "locations_quantities_adverbs"))) %>%
  tidyboot::tidyboot_mean(d, na.rm=T)

prod_cat %>% ggplot(aes(x=reorder(language, mean), y=mean)) + geom_point(alpha=.7) +
  geom_linerange(aes(ymin=ci_lower, ymax=ci_upper), alpha=.7) +
  facet_wrap(. ~ category) + coord_flip() + 
  theme_classic() + ylab("Mean item easiness") + xlab("Language")
#ggsave("xling_diff_by_category_WSprod.pdf", width=10, height=16)
```


## Cross-linguistic similarities 

We look at the Spearman correlation between the item difficulty of each language compared to each other language. 
We might expect this to recapitulate the historical relationship between languages, with more similar languages having more similar item difficulties (e.g., Quebecois and European French).

```{r xling-sim, fig.width=10, fig.height=10, set.cap.width=T, num.cols.cap=2, fig.cap = "Cross-linguistic similarity (Spearman correlation) of IRT item difficulty from the CDI:WS."}
# parm diffs in low- vs. high-data langs
#t.test(xldf$d, xldf_lowd$d) # -.45 vs. -.02 (closer to prior of 0 in low-data langs)
#t.test(xldf$a1, xldf_lowd$a1) # higher discrim in high-data langs

tmp <- get_cross_ling_difficulty_cors(xldf)
prod_cors = tmp$prod_cors
prod_sims = tmp$prod_sims

Colors = brewer.pal(11,"Spectral")
diag(prod_cors) = NA

#pdf(file="xling_difficulty_similarity_WSprod.pdf", width=13, height=13)
heatmap.2(prod_cors, col=Colors, margins=c(10,10))
#dev.off()

# compare to one of these?
# http://www.elinguistics.net/
# https://www.ezglot.com/most-similar-languages.php
# 28 x 28 = 784 values...scrape?

```


## Evaluating Thresholds for Inclusion on Swadesh CDI

It is unclear how many existing forms a uni-lemma should be included on in order for it to be considered for inclusion on a Swadesh CDI. 
Instead of setting an arbitrary threshold, we chose to try all possible values of $k$ (i.e., number of forms on which a uni-lemma is included, $k \in \{2,..,26\}$), and chose the $k$ which yields the set of Swadesh CDI (S-CDI) items with the highest total test information.
The table below shows test information for each threshold, with the highest S-CDI test information achieved by $k=9$ (66577).

```{r, echo=F}
load(here("data/Swadesh_grid_search_final.Rdata"))

swad_tab <- swad_agg %>% filter(sublist=="Swadesh") %>%
  filter(language!="French (French)") %>% # test info is WAY high -- something fishy
  mutate(k = as.factor(k)) %>%
  group_by(k, N, mean_diff, mean_sd, sd_sd, sd_thresh, swad_n, swad_mean_diff, swad_diff_sd) %>%
    summarise(N_avail = mean(N_avail),
              `S-CDI vs CDI:WS` = mean(sumscore_cor), # want sd(sumscore_cor) and sd(test_info) ?
              `S-CDI Info` = round(mean(test_info),0))

rand_tab <- swad_agg %>% filter(sublist=="random") %>%
  filter(language!="French (French)") %>% 
  mutate(k = as.factor(k)) %>%
  group_by(k) %>%
    summarise(`Rand vs CDI:WS` = mean(sumscore_cor), # want sd(sumscore_cor) and sd(test_info) ?
              `Rand Info` = round(mean(test_info),0))

grid_tab <- left_join(swad_tab, rand_tab) %>%
  relocate(`Rand vs CDI:WS`, .after=`S-CDI vs CDI:WS`) %>%
  relocate(`Rand Info`, .after=`S-CDI Info`)

grid_tab %>% select(-swad_diff_sd) %>%
  #mutate(d_diff = mean_diff - swad_mean_diff,
  #       info_diff = round(`S-CDI Info` - `Rand Info`,0)) %>%
  rename(`Mean d` = mean_diff, 
         `S-CDI pool` = swad_n,
         `S-CDI Mean d` = swad_mean_diff,
         `S-CDI per form` = N_avail) %>%
  kableExtra::kable(digits=3) %>%
    html_table_width(rep(72, ncol(swad_agg)+2))
```


## Swadesh CDI vs. Full CDI:WS

For the 26 languages that the IRT models were trained on:

```{r, results='asis'}
load(here("data/Swadesh_comparisons.Rdata"))

#xx_tab <- xtable::xtable(xx, digits=c(3), 
#                       caption = "Swadesh CDI vs. Full CDI:WS scores for 26 training languages.")
#print(xx_tab, type="html", comment = F, table.placement = "H", include.rownames=FALSE)

xx %>% kable(digits=3, caption="Swadesh CDI vs. Full CDI:WS scores for 26 training languages.") %>%
  html_table_width(rep(70, ncol(xx)))
```

### Generalization Test

Comparison to the eight low-data languages, with the 10 additional difficult words added.
Note that many proposed Swadesh items are not actually on the CDI:WS forms available to test generalization. 

```{r, results='asis'}
#gen_tab <- xtable::xtable(gen_xx_ext, digits=c(3), 
#                       caption = "Swadesh CDI with difficult words vs. Full CDI:WS scores for 8 generalization languages.")
#print(gen_ext_tab, type="html", comment = F, table.placement = "H",
#      include.rownames=FALSE)

gen_xx %>% kable(digits=3, caption="Swadesh CDI with difficult words vs. Full CDI:WS scores for 8 generalization languages.") %>%
  html_table_width(rep(70, ncol(gen_xx)))
```


## Swadesh-CDI Items

```{r}
good_prod <- swad_lists[[9]]
#kableExtra::kable(good_prod, digits=2)
good_prod %>% DT::datatable() %>%
  DT::formatRound(columns=4:24, digits=3)
```

