---
title: "Measuring Children's Early Vocabulary in Low-Resource Languages Using a Swadesh-style Word List"
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf George Kachergis* (kachergis@stanford.edu)} 
    \AND {\large \bf Alvin Wei Ming Tan* (tanawm@stanford.edu)}
    \AND {\large \bf Virginia A. Marchman (marchman@stanford.edu)} 
    \AND {\large \bf Michael C. Frank (mcfrank@stanford.edu)} \\ Department of Psychology, 450 Jane Stanford Way \\ Stanford, CA 94305 USA}

abstract: >
    Early language skill is predictive of later life outcomes, and is thus of great interest to
    developmental psychologists and clinicians. The Communicative Development Inventories (CDI),
    a parent-reported inventory of early-learned vocabulary items, have proven to be valid 
    and reliable instruments for measuring children's early language skill. CDIs have been
    painstakingly adapted to dozens of languages, and cross-linguistic comparisons thus far show both 
    consistency and variability in language acquisition trajectories. However, thousands of languages do 
    not yet have CDIs, posing a significant barrier to increasing the diversity of languages that are 
    studied. Here, we propose a method for selecting candidate words to include on new CDIs, leveraging
    analysis of psychometric properties of translation-equivalent concepts that are frequently included on 
    existing CDIs. Leveraging 26 datasets from existing CDIs, we propose a list of 229 concepts that have
    low variability in their cross-linguistic learning difficulty. This pool of common concepts---analogous 
    to the "Swadesh" lists used in glottochronology---can be used as a starting point for future 
    CDI adaptations. We test how well the proposed list generalizes to data from 8 additional languages.
    
keywords: >
    early language learning; CDI; psychometrics; cross-linguistic comparison; Swadesh vocabulary
    
output: cogsci2016::cogsci_paper
#final-submission: \cogscifinalcopy
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, 
                      fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, 
                      message=F, sanitize = T)
```

```{r, libraries}
library(png)
library(grid)
library(xtable)
require(mirt)
require(tidyverse)
require(ggpubr)
require(tidyboot)
require(here)
library(patchwork)

source(here("cross-ling-comparison-helpers.R"))

set.seed(42)
```

# Introduction

Tools that enable valid assessments of children’s early language abilities are invaluable for researchers, clinicians and parents, as early language skill is predictive of educational outcomes years later [e.g., @bleses2016]. 
The MacArthur-Bates Communicative Development Inventories [CDIs, @Fenson2007; @marchman2023] are parent report assessments that provide reliable and valid estimates of children's early vocabulary size and other aspects of early communicative development, such as use of gestures and of word combinations. 
Parent report is a relatively quick and low-cost method to assess early language skills as it takes advantage of the fact that parents are "natural observers" of their child's skills and does not depend on a child engaging with an unfamiliar experimenter.

Over the years, the CDIs have been adapted to dozens of languages, with forms now available in English, Spanish, French, Hebrew, and Mandarin, to name just a few.
Recently, data from more than 85,000 CDIs in 38 languages have been archived in a central repository [Wordbank, @frank2017]. 
These data have revealed both cross-linguistic consistency and variability in early language skills, with insights from these patterns informing theories of early language learning [@frank2021]. 
For example, cross-linguistic analyses indicate that measures of vocabulary size are tightly correlated with other aspects of early language skill, like gesture and grammatical competence. 
Thus, over development, the language system is "tightly woven" [@bates1994; @frank2021] and early vocabulary size serves as a good proxy measure of children's overall language skill.

On the CDIs, vocabulary size is assessed via a checklist format, which enables caregivers to scan and recognize words their child produces or understands, rather than relying on recall alone. 
For example, the American English CDI Words & Sentences (CDI:WS) form, targeting children 16-30 months of age, is comprised of 680 words from 22 semantic categories, including nouns (e.g., Body Parts, Toys, and Clothing), action words, descriptive words, and closed-class words such as pronouns. 
Items on this original CDI:WS were chosen to reflect a range of difficulty levels (i.e., easy, moderate, and more difficult), as well as capture the linguistic and societal contexts of (most) children living in the US.
Short versions of the CDI:WS forms are also available [e.g., @Fenson2000], consisting of a set of $\sim100$ items that generate scores that more strongly correlate with scores on the long forms, while retaining representation across a broad set of semantic categories.

<!-- gap! -->
Creating a new CDI requires a lot of effort and resources, presenting a daunting barrier to increasing the diversity of languages studied.
Following the guidelines[^1] from the MacArthur-Bates CDI Advisory Board, the process of adapting a CDI for a language other than American English goes well beyond simply translating items on these forms to that new language. 
While the process can begin with identifying translation equivalents (i.e., items that capture the same general concept in both languages, e.g., "dog" in English, and "perro" in Spanish), the final item set must then be filtered so that all items appropriately reflect the linguistic and sociocultural context of the children learning that language. 
This process usually requires considerable time and effort by researchers who are both native speakers of the language and who have experience with children, to first select and identify translation equivalents and to then iteratively add, refine, and pilot the new CDI in the target language. 
Because the goal is to obtain the set of items that best capture general trends and individual differences in that language, the items across CDIs in different languages do not necessarily overlap to a great extent. 
For example, the American English CDI:WS and Mexican Spanish CDI:WS forms each have 680 words, but only have 463 overlapping concepts (68%).

[^1]: [https://mb-cdi.stanford.edu/adaptations.htm](https://mb-cdi.stanford.edu/adaptations.htm)


It is well-established that, all over the world, early-learned words reflect the people and things that children are likely to experience, that is, words for family members, animals, and common household objects [@tardif2008baby; @frank2021]. 
Given this finding, it is reasonable to ask: Is there a single set of translation equivalents that would meet the criteria for inclusion on CDIs from multiple languages? 

To facilitate this effort, it is useful to leverage Item-Response Theory [IRT, @embretson2013] models. 
IRT models infer both the abilities of test takers and the difficulty of individual test items (i.e., words), along standardized dimensions. 
Recent work using IRT models has facilitated our understanding of the psychometric properties of specific CDI instruments. 
As such, they offer the potential to not only yield more accurate measures of children's language ability, but also to enable the construction of language-specific Computerized Adaptive Tests (CATs), which choose the next test item based on the responses to the previous items, and thus quickly hone in on the test taker's language ability. 
CAT-based CDIs presenting 50 or fewer items have been found to strongly correlate with scores on the full CDI:WS [@chai2020;@mayor2019;@Makransky2016]. 
A general method for creating CDI CATs that work well across a broader age range (12--36 months) has been proposed, and tested for American English and Mexican Spanish [@Kachergis2022]. 
However, the IRT model driving each CAT needs to be trained on a large and normative dataset, which may not be available in a given language. <!-- define normative? i.e., representative of a target population -->
To date, the IRT models are fitted separately for each language, and the fitted parameters (e.g., word difficulty) are likely to vary across languages.

The goal of the current study is to use IRT modeling in conjunction with data from Wordbank to examine whether there might be a core set of concepts that are frequently included on CDIs, and---importantly---whether enough of them are of roughly equal difficulty across many languages to allow them to be used as candidate items in new languages. 
This work takes its inspiration from the fields of lexicostatistics and glottochronology, where researchers [notably, @Swadesh1971] have proposed lists of common concepts that exist in all catalogued languages, in order to quantify the genealogical relatedness and dates of divergence of languages. 
For example, the original Swadesh list contains 100 words, comprised of categories including common pronouns (_I_, _you_, _we_), animals (_man_, _fish_, _bird_, _dog_), objects (_tree_, _leaf_, _sun_, _mountain_), and verbs (_die_, _see_, _sleep_, _kill_). 
Extending this work to the development of a universal CDI, or “Swadesh CDI,” would include many of the concepts that researchers have chosen to include on several CDI:WS adaptations, and which have relatively similar difficulty across many languages. 
If such a list were generalizable to other languages, it could serve as a helpful starting point for the development of new CDI adaptations, since the constituent words would already have good cross-linguistic difficulty estimates.

<!--More broadly, this work examines which types of words (and their corresponding concepts) are more or less similar in terms of the ease with which they are learned across languages, revealing commonalities as well as idiosyncrasies in children’s early experiences. 
We can ask: Which semantic categories, e.g., animals, household objects, food and drink, or another category, are most consistently learned across languages? Which are more variable? 
These types of cross-linguistic comparisons may give new insight into theoretical questions surrounding the similarity and differences between language experience and development in different cultural and linguistic contexts.--> <!--AT: this para seems to not be the desired outcome of the S-CDI so I removed it.-->

In particular, our contributions are 1) to revise and extend a set of translation-equivalent concepts in Wordbank, 2) to fit IRT models to 26 CDI:WS datasets, 3) to evaluate 25 candidate lists of Swadesh CDI items from a cross-linguistic comparison of concept difficulty and inclusion, 4) to identify and characterize the most informative Swadesh CDI list of 229 concepts, and 5) to test its generalization to a set of eight additional low-data languages. 
We then make a concrete proposal for how this Swadesh CDI list could be used to create future CDI adaptations, expanding the diversity of languages studied. 
We end by discussing the strengths and weaknesses of our approach.
<!-- Our full analysis, the Swadesh CDI list, and other information valuable for developing a new CDI are openly available on [OSF](https://osf.io/8swhb/?view_only=6f6ab9818f2a4bb288e05ca9e12f540c). -->

# Methods

## Item Response Theory

A variety of IRT models targeting different types of testing scenarios have been proposed [see @Baker2001 for an overview], but for the dichotomous responses that parents make for each item (word) regarding whether their child can produce a given word, we used the popular 2-parameter logistic (2PL) model that is best justified for CDI data out of four standard models [see @Kachergis2022].

The 2PL model jointly estimates for each child $j$ a latent ability $\theta_j$ (here, language skill), and for each item $i$ two parameters: the item's difficulty $b_i$ and discrimination $a_i$, described below.
In the 2PL model, the probability of child $j$ producing a given item $i$ is 
 
 $$P_{i}(x_i = 1 | b_{i},a_{i},\theta_j ) = \frac{1}{1 + e^{-D a_{i}(\theta_j - b_i )}}$$

where $D$ is a scaling parameter ($D=1.702$) which makes the logistic more closely match the ogive function used in a standard factor analysis [@R-mirt; @reckase2009].
Children with high latent ability ($\theta$) will be more likely to produce any given item than children with lower latent ability, and more difficult items will be produced by fewer children (at any given $\theta$) than easier items.
The discrimination ($a_i$) adjusts the slope of the logistic (in the classic 1-parameter logistic "1PL" model, the slope is always 1). 
Items with higher discrimination (i.e., slopes) better distinguish children above vs. below that item's difficulty level, and hence are generally more useful.
While other standard IRT models exist (e.g., the 3PL model adds a "guessing" parameter for each test item), a recent study found the 2PL model most appropriate for multiple Wordbank datasets [@Kachergis2022].

## Datasets

```{r, load-wg-data, eval=F, echo=F}
# WG production fits
load(here("data/multiling_2pl_WG_prod_fits.Rdata"))
wg_coefs = coefs
languages = names(coefs)
wg_tab <- get_item_n_subject_counts(models)
#sort(languages)

#load(here("data/xling_WG_abilities.Rdata"))
load(here("data/xling-WGprod-IRTparms.Rdata"))
# 24 languages, 9611 parameters
wg_parms <- xldf
```

```{r, load-ws-data, echo=F, results="asis"}
# WS production fits
load(here("data/multiling_2pl_WS_prod_fits.Rdata"))
load(here("data/xling-WSprod-IRTparms.Rdata")) # 23179, 

xldf <- xldf %>% mutate(uni_lemma = ifelse(uni_lemma=="NA", NA, uni_lemma)) %>%
  mutate(d = -d) %>% # easiness -> difficulty
  filter(!is.na(d)) # removes 1 Slovak word with NA parameters (because all responses were 0)

ws_tab <- get_item_n_subject_counts(models)

# WG langs that don't have WS: British Sign Language, Spanish (Chilean), American Sign Language, English (British)

it_tab <- ws_tab %>% #rename(`WS items`=items, `WS N`=N) %>%
  #left_join(wg_tab %>% rename(`WG items`=items, `WG N`=N)) %>%
  arrange(desc(`N`))

languages = subset(ws_tab, N>=200)$Language
low_data_langs <- it_tab %>% filter(`N`<200)

bold <- function(x) {paste('{\\textbf{',x,'}}', sep ='')}

tab1 <- xtable::xtable(it_tab, digits=c(0), 
                       caption = "CDI:WS items and subjects (N) per dataset. The final 8 datasets were used for a generalization test.")
print(tab1, type="latex", comment = F, 
      sanitize.colnames.function=bold, # bold the header row
      table.placement = "H", hline.after=26,
      include.rownames=FALSE, size="\\fontsize{9pt}{10pt}\\selectfont")
```


We report IRT analyses for twenty-six languages from Wordbank [@frank2017], comprising production data from CDI:WS vocabulary checklists that have at least 200 administrations.[^2]
Data from the first twenty-six rows of Table 1 (Norwegian through Dutch) were used to select a pool of words with approximately equal cross-linguistic difficulty.
CDI:WS production data from an additional eight languages (bottom of Table 1: Greek through Persian) had too few participants to be analyzed with IRT.
Datasets for these languages were used to test how well the selected pool can be expected to generalize to new languages.

### Uni-lemmas

Comparison across languages requires a method to map between words that correspond to broadly similar concepts across languages. 
As such, each item on the CDI:WS for each language was mapped onto a set of “universal lemmas” or “uni-lemmas”, which are approximate cross-linguistic conceptual mappings of words. 
For example, “chat” (French) and “gato” (Spanish) both correspond to the same uni-lemma, _cat_. 
These mappings were recently updated to improve their quality and systematicity, and to increase coverage across items and languages. 
This new set of uni-lemmas was constructed based on glosses provided by the original contributors of the Wordbank datasets, which were then verified by native or advanced proficient speakers of the language, and cleaned to increase their consistency across languages. 
All uni-lemmas are accessible from Wordbank; details about the recent update can be found at [https://github.com/langcog/update_unilemmas](https://github.com/langcog/update_unilemmas).

### Participants


```{r participant-data, eval=F, echo=F}
demo <- tibble()
for(lang in languages) {
  load(here(paste("data/WS/",lang,"_WS_data.Rdata", sep='')))
  demo <- bind_rows(demo, d_demo %>% 
                      select(data_id, age, production, language))
}

save(demo, file=here("data/WS_combined_demo.Rdata"))
```

```{r combined-demo-data, echo=F}
load(here("data/WS_combined_demo.Rdata"))
# cor.test(demo$age, demo$production, na.rm=T) # r=.70

# 45402 in the 26 high-data languages
demo_age <- demo %>% group_by(language) %>%
  summarise(n = n(), age = mean(age))

#demo %>% 
#  ggplot(aes(x=age, fill=language, color=language)) + 
#    geom_histogram(position="identity", alpha=0.2)
```


The Wordbank datasets consisted of the CDI:WS production data for `r sum(ws_tab$N)` children aged 16--30 months on `r sum(ws_tab$items)` items across `r nrow(ws_tab)` forms.
Note that the distributions of demographic variables (age, sex, etc.) of these datasets are not matched, so comparing overall language ability estimates across languages would be ill-advised. 
(See @frank2021 for a discussion of effects of demographic variables on vocabulary development.) 
Thus, we focused only on the estimated item parameters, and in particular the variability of item difficulty ($b_i$).

[^2]: [http://wordbank.stanford.edu/contributors](http://wordbank.stanford.edu/contributors)



### Instruments

<!-- When a CDI:WG form was administered, caregivers were asked to indicate for each vocabulary item whether their child 1) understands that word ("comprehends") or 2) both understands and says ("produces") that word. 
Leaving the item blank indicates that the child neither comprehends nor produces that word. -->
When a CDI:WS forms was administered, caregivers were asked to indicate for each vocabulary item on the instrument whether or not their child can recognizably produce (say) the given word in an appropriate context.

"Produces" responses were coded as 1 and all other responses were coded as 0.
Our datasets consisted of a dichotomous-valued response matrix for each language, of size $N$ subjects $\times$ $W$ words.
All models, data, and code for reproducing this paper are available on OSF[^3].

# Results

```{r}
xldf <- xldf %>% filter(!is.na(uni_lemma)) # 21824 out of 23020 have uni-lemmas defined 95%
xldf_lowd <- xldf %>% filter(is.element(language, low_data_langs$Language)) # 5464
xldf <- xldf %>% filter(!is.element(language, low_data_langs$Language)) # 17715

# align some categories: Russian has descriptive adjectives and adverbs
xldf[which(xldf$category=="descriptive_words (adjectives)"),]$category = "descriptive_words"
xldf[which(xldf$category=="outside_places"),]$category = "outside"
```


```{r, include=F}
# 1839
uni_ag <- xldf %>% filter(uni_lemma!="NA") %>%
  group_by(uni_lemma) %>% # category <- but many uni-lemmas are categorized differently across languages
  summarise(d_m=mean(d), a1_m=mean(a1), d_sd = sd(d), a1_sd = sd(a1), n=n())

# The more times a uni-lemma appears, the easier it tends to be:
cor.test(uni_ag$d_m, uni_ag$n)
#plot(uni_ag$d_m, uni_ag$n)

pval_str <- function(pval) {
  p_str = ifelse(pval<.001, "<.001", 
                ifelse(pval<.01, "<.01", 
                       ifelse(pval<.05, "<.05", 
                              paste0("=",round(pval, 2)))))
  return(p_str)
}

# given a cor.test(), return the in-line reporting, e.g. "$r=-0.37$, $t(1950)=-17.64$, $p<.001$" )
report_cor <- function(ctest) {
  pval = pval_str(ctest$p.value)
  paste0("$r=",round(ctest$estimate,2),"$, $t(",ctest$parameter,")=",round(ctest$statistic, 2),"$, $p",pval,"$")
}

report_ttest <- function(ttest) {
  pval = pval_str(ttest$p.value)
  paste0("$t(",round(ttest$parameter,0),")=",round(ttest$statistic, 2),"$, $p",pval,"$")
}

# variability in difficulty also correlated number of forms uni-lemma is on

# 990 singletons
uni1 <- uni_ag %>% filter(n==1) 
mean(uni1$d_m) # 1.38

uni_ag_gt1 <- uni_ag %>% filter(n>1) # 1311

# singletons are harder than unis appearing on multiple forms?
report_ttest(t.test(uni1$d_m, uni_ag_gt1$d_m))

# no cor bw difficulty and variability -- good
report_cor(cor.test(uni_ag_gt1$d_m, uni_ag_gt1$d_sd))
```

Across the 26 IRT models for different CDI:WS forms, difficulty and discrimination parameters for a total of 23020 items were fitted.
Of those items, 95% had uni-lemmas defined, with a median of 693 per CDI:WS form (range: 553 in Czech to 804 in Cantonese). 
A total of `r length(unique(xldf$uni_lemma))` uni-lemmas were defined across the 26 CDI:WS forms, but `r nrow(uni1)` of these were singletons, appearing on only one of the 26 forms.\footnote{These singletons were significantly more difficult than the `r nrow(uni_ag_gt1)` uni-lemmas appearing more than once ($M_1=$ `r round(mean(uni1$d_m),2)`; $M_{>1}=$ `r round(mean(uni_ag_gt1$d_m),2)`; `r report_ttest(t.test(uni1$d_m, uni_ag_gt1$d_m))`).} 
There was a significant relation between how often a uni-lemma appears and its difficulty: the more often a uni-lemma appears, the _easier_ it tended to be (`r report_cor(cor.test(uni_ag$d_m, uni_ag$n))`).
Moreover, there was a weak but significant relation between the number of forms a uni-lemma appears on and its cross-linguistic variability (`r report_cor(cor.test(uni_ag$d_sd, uni_ag$n))`).
It is perhaps intuitive that lower-variability items tend to be earlier-learned, and are thus often selected to be on CDI forms, echoing prior work characterizing the consistency of children's first words across several languages [@tardif2008baby].
However, these modest but significant correlations were also important to keep in mind as we chose our Swadesh CDI candidates, as selecting too many easy items could result in older children being at ceiling.

<!--Fortunately, among the uni-lemmas appearing on at least two forms, there is no significant relation between the uni-lemmas' difficulty and cross-linguistic variability in difficulty.-->

```{r fig-mean-item-diff, include=F, fig.cap=c("Mean item difficulty for each CDI:WS form. Bars represent bootstrapped 95\\% confidence intervals. Color shows mean age (months) of the sample, which is correlated with mean word difficulty (r=.57)."), fig.height=4.3, fig.width=3.4}
# re: Philip's question of are item difficulties comparable across languages:
lang_d_m <- xldf %>% filter(!is.na(d)) %>% # one Slovak word has NA parms..
  group_by(language) %>% 
  tidyboot::tidyboot_mean(d, na.rm=T) %>%
  arrange(desc(mean))

lang_d_m <- lang_d_m %>% left_join(demo_age %>% select(language, age))
# problem: average item difficulty per language is strongly related to mean sample age
# cor.test(lang_d_m$mean, lang_d_m$age) # r=.57  t(24) = 3.44, p = 0.002

lang_d_m %>% ggplot(aes(x=reorder(language, -mean), y=mean)) + coord_flip() + 
  geom_point(aes(color=age)) + theme_classic() +
  geom_linerange(aes(ymin=ci_lower, ymax=ci_upper), alpha=.7) +
  theme(legend.position="bottom") +
  xlab("Language") + ylab("Mean item difficulty")
```


[^3]: OSF repository: [https://osf.io/8swhb/](https://osf.io/8swhb/?view_only=6f6ab9818f2a4bb288e05ca9e12f540c).


## Identifying Swadesh CDI Candidates

The goal was to choose uni-lemmas with low variability in their cross-linguistic difficulty---that is, uni-lemmas that are similarly difficult to learn across languages.
There were two thresholds that had to be defined to select Swadesh CDI (S-CDI) candidates: 1) $k$, the number of current CDI:WS forms a uni-lemma must appear on, and 2) $v_{min}$, the threshold for variability in cross-linguistic difficulty. 
For $v_{min}$, we simply considered the uni-lemmas with a variation in cross-linguistic difficulty that was half a standard deviation less than average, among the set of uni-lemmas that were on at least $k$ forms. 
Assuming variability was normally-distributed, this criterion preserved the least-variable 19% of the uni-lemmas. <!-- 19.1% of a normal distro is <(mean-.5SD)  -->

It was less clear how to choose $k$, the minimum number of forms a uni-lemma must appear on to be considered eligible.
As reported above, the more forms a uni-lemma appeared on, the easier that concept tended to be, and the greater its cross-linguistic variability in difficulty ($SD(d)$) tended to be.
Thus, the higher the threshold $k$, the more bias there would be toward the Swadesh candidates being easier than typical, as well as more variable in difficulty.
Having a Swadesh CDI list that was much easier than the CDI:WS is undesirable, as it may show ceiling effects for older children, while selecting words that are more variable in difficulty may result in a test that generalizes less well to other languages.
Of course, the higher $k$, the fewer uni-lemmas there were available to choose from: 1303 uni-lemmas appeared on 2 or more CDIs, but only 30 appeared on all 26 CDIs.

Because choosing $k$ is difficult _a priori_, we opted to do a grid search across all possible values. 
For each $k$, we found the set of $N$ uni-lemmas appearing $k$ times on each form, calculated the mean and standard deviation of the difficulties of those items, and then identified the subset of those $N$ items with $SD(d) < v_{min}$: these comprised the $k^{\textrm{th}}$ Swadesh list, as shown in Figure 1 (top).
For each candidate Swadesh list, we evaluated the total test information yielded by that Swadesh list for each of the 26 CDI:WS forms. 
Total test information was calculated as the integral across a range of test-taker abilities (here, $\theta \in \{-4,4\}$) of the information yielded on a test, calculated using the IRT difficulty and discrimination parameters for those test items.
Finding the $k$ that yielded the highest total test information was the main objective (see online Appendix for details).

As shown in Figure 1 (bottom), test information was maximized with $k=9$, and from these 1309 uni-lemmas appearing on at least 9 forms, 229 Swadesh-CDI items were selected for having cross-linguistic variability below the threshold.
The mean difficulty of the S-CDI items was also quite close to the difficulty of the other uni-lemmas that appeared on at least 9 forms ($M(d_{Swad})=0.33$; $M(d_{other})=0.37$), suggesting that the S-CDI items were fairly representative.



```{r, echo=F, include=F}
# problem: some languages have multiple items matching a single uni_lemma
# (e.g., Croatian has stric=uncle and ujak=uncle; Turkish has 3 "market"s )
prod_pars <-
  xldf %>% arrange(desc(language), desc(uni_lemma), desc(a1)) %>% # get most discriminating uni_lemma per lang
  #filter(is.element(uni_lemma, names(short_list))) %>%
  select(uni_lemma, category, lexical_category, language, d) %>%
  group_by(uni_lemma, language) %>%
  slice(1) %>% 
  pivot_wider(names_from = language, values_from = d)

# only want 1 
uni_per_form <- xldf %>% arrange(desc(language), desc(uni_lemma), desc(a1)) %>% # get most discriminating uni_lemma per lang
  select(uni_lemma, category, lexical_category, language, d) %>%
  group_by(uni_lemma, language) %>%
  slice(1) %>%
  group_by(uni_lemma) %>%
  summarise(d_m = mean(d), d_sd = sd(d), n = n()) %>%
  filter(!is.na(d_m)) # one uni-lemma: "look for"

uni_per_form %>% ggplot(aes(x=jitter(n), y=d_sd, color=d_m)) + 
  geom_point(alpha=.3) + theme_bw()
```



```{r grid-search, eval=F}
# how many Swadesh words when looking at all uni-lemmas included in at least k languages?
# (same mean(sd) - sd(sd) threshold applied to all, but could make that a grid, e.g. sd_thresh = seq(0, 2, .25)
swad_lists <- list()
swad_agg <- tibble()

# has to be at least 2 forms, as SD only defined for 2+
for(k in 2:26) {
  uni_k <- uni_per_form %>%
    filter(n >= k)
  N = nrow(uni_k)
  mean_diff = mean(uni_k$d_m) # average difficulty of uni-lemmas on at least k forms
  mean_sd = mean(uni_k$d_sd) # average SD of difficulty of uni-lemmas on at least k forms
  sd_sd = sd(uni_k$d_sd) # variability of SD(diff)
  sd_thresh = mean_sd - .5*sd_sd
  swad_lists[[k]] = uni_k %>% filter(d_sd < sd_thresh)
  
  xx_g <- run_comparisons(xldf, languages |> sort(), swad_lists[[k]]$uni_lemma, 
                        rand_method = "items", rand_comparisons = 1000,
                        metrics = c("sumscore_cor", "test_info")) # theta_cor breaks on French (French): incompatible dimensions
  
  xx_agg <- xx_g |> 
    group_by(language, sublist, metric) |> 
    summarise(mean = mean(value),
              N_avail = mean(N),
              sublist_mean_diff = mean(mean_d)) |> 
    spread(metric, mean) |>
    mutate(k = k, 
           N = N,
           mean_diff = mean_diff,
           mean_sd = mean_sd,
           sd_sd = sd_sd,
           sd_thresh = sd_thresh,
           swad_mean_diff = mean(swad_lists[[k]]$d_m),
           swad_diff_sd = sd(swad_lists[[k]]$d_m),
           swad_n = nrow(swad_lists[[k]]),
           sublist_mean_diff = sublist_mean_diff)
  
  swad_agg <- bind_rows(swad_agg, xx_agg)
}



# which items crop up over and over? (have to decide which items to discuss in the paper!)
all_swad <- c()
for(k in 2:26) {
  all_swad = c(all_swad, swad_lists[[k]]$uni_lemma)
}
swad_freq = sort(table(all_swad), decr=T)

save(swad_agg, swad_lists, swad_freq, file=here("data/Swadesh_grid_search_final.Rdata"))

#library(wordcloud)
#library(RColorBrewer)
#wordcloud(words = names(swad_freq), freq = swad_freq, min.freq = 1,           
#          max.words=100, random.order=FALSE, rot.per=0.35, scale=c(3.5,0.25),
#          colors=brewer.pal(8, "Dark2"))

library(wordcloud2)
wordcloud2(data=data.frame(word = names(swad_freq[1:150]), 
                           freq = as.vector(swad_freq[1:150]-13)), size=.5, color='random-dark')
```

```{r load-swadesh-lists, echo=F}
load(here("data/Swadesh_grid_search_final.Rdata"))


# looking at appendix table, k=9 seems the best (highest test info, v close to same full correlation as rand)
good_prod = swad_lists[[9]]

en_good_prod <- good_prod %>% left_join(xldf %>% filter(language=="English (American)"))
# how many Swadesh items not on English?
not_en_good <- subset(en_good_prod, is.na(item_id)) # 28 
#not_en_good$uni_lemma

```


```{r, include=F}
# production sumscore vs. Swadesh prod sumscore
all_unis <- bind_rows(xldf, xldf_lowd) 

get_sumscores_vs_swad <- function(all_unis, languages, swad_list, form="WS") {
  xx <- tibble()
  for(lang in languages) {
    load(here(paste("data/",form,"/",lang,"_",form,"_data.Rdata", sep='')))
    swad_l <- subset(all_unis, language==lang & is.element(uni_lemma, swad_list)) 
    swad_cor = cor(rowSums(d_prod, na.rm=T), rowSums(d_prod[,swad_l$item_id], na.rm=T))
    tmp <- tibble(WS_score=rowSums(d_prod, na.rm=T), 
                  swad_score=rowSums(d_prod[,swad_l$item_id], na.rm=T)) %>% 
      mutate(language = lang, Nswad = nrow(swad_l), Nws = ncol(d_prod))
    if(nrow(d_demo)==nrow(tmp)) {
      tmp$age = d_demo$age
    } else { # e.g., Spanish (Mexican) has a d_demo dim mismatch, so can't easily align
      tmp$age = NA 
    }
    xx <- xx %>% bind_rows(tmp)
  }
  return(xx)
}

gen_df <- get_sumscores_vs_swad(all_unis, low_data_langs$Language, good_prod$uni_lemma, form="WS") %>%
  mutate(prop_swad = swad_score / Nswad,
         prop_ws = WS_score / Nws) 

gen_ag <- gen_df %>% group_by(language) %>%
  summarise(swad_cor = cor(prop_swad, prop_ws),
            Nws = median(Nws),
            Nswad = median(Nswad))

mean(gen_ag$Nswad)
# only 153 of the 230 Swadesh items are on each low-data lang's WS, on average
mean(gen_ag$swad_cor) # .990 mean cor
cor.test(gen_ag$swad_cor, gen_ag$Nswad) # the more Swadesh items, the better the cor
# t = 2.8013, df = 6, p-value = 0.03111

gen_df %>%
  ggplot(aes(x=prop_swad, y=prop_ws, color=age)) + geom_point(alpha=.5) + theme_classic() + 
  geom_smooth() + ylab("Proportion of Known CDI:WS Items") +
  xlab("Proportion of Known Swadesh CDI Items") +
  ggtitle("Generalization Test")

```



```{r grid-search-plots, fig.cap=c("Number of uni-lemmas appearing on at least $k$ lists (top). Total test information for Swadesh lists, and random tests of the same length (bottom)."), fig.height=4.5, fig.width=3.4}
plot_scale <- scale_color_manual(breaks = c("full", "Swadesh", "random"),
                     values = c("black", "#f8766d", "#00bfc4"))

plot_k_n <- ggplot(data = swad_agg |> select(k, N, swad_n) |> distinct() |> 
                     pivot_longer(cols = c("N", "swad_n"), names_to = "sublist", values_to = "N") |> 
                     mutate(sublist = sublist |> factor(levels = c("N", "swad_n", "r"),
                                                        labels = c("full", "Swadesh", "random"))) |> 
                     bind_rows(tibble(k = 0, sublist = "random", N = 0))) +
  geom_line(aes(x = k, y = N, col = sublist)) +
  # geom_line(aes(x = k, y = N)) +
  # geom_line(aes(x = k, y = swad_n), color="#00bfc4", linetype="dashed") + 
  #geom_text(x=5, y=1000, label="All uni-lemmas on k or more forms", color="black") + 
  # geom_text(x=10, y=370, label="Swadesh lists", color="blue") + 
  theme_classic() +
  plot_scale +
  scale_x_continuous(breaks = c(5,10,15,20,25)) + 
  labs(y = "Number of uni-lemmas",
       col = "Word list") +
  coord_cartesian(xlim = c(2, 26))

plot_k_testinfo <- ggplot(data = swad_agg |> group_by(sublist, k) |> summarise(test_info = mean(test_info))) +
  geom_line(aes(x = k, y = test_info, col = sublist)) +
  geom_vline(aes(xintercept=9), linetype="dashed") + 
  theme_classic() + theme(legend.position="bottom") + 
  plot_scale +
  scale_x_continuous(breaks = c(5,10,15,20,25)) + 
  labs(y = "Total test information")

ggarrange(plot_k_n, plot_k_testinfo, nrow=2, common.legend = TRUE, legend="bottom")
```

## Characterizing the Swadesh-CDI 

```{r, out.width="\\linewidth", include=TRUE, fig.align="center", fig.cap=c("The 229 Swadesh CDI concepts by semantic category, and 10-item extension (in red). Italics denote the 28 items not included on the American English CDI:WS."), echo=FALSE}
knitr::include_graphics("figs/SwadeshCDI_list.pdf")
```

The `r nrow(good_prod)` S-CDI items, shown in Fig. 3, represented 21 of the 22 semantic categories present on the original American English CDI:WS form, with concrete nouns, action words, and adjectives being most prevalent, and connecting words, quantifiers, helping verbs, and articles being rare, and question words being unrepresented (but see proposed extension below).
54% of the S-CDI concepts were nouns, 22% were predicates (verbs and adjectives), 17% were function words, and 7% belonged to other lexical categories.
This breakdown was comparable to the lexical category percentages on the 680-item English CDI:WS (46% nouns, 24% predicates, 15% function words, and 5% other), suggesting that the S-CDI list did not show a particular lexical category bias.
The S-CDI items were also present on more forms than typical in the selection set: on average, each item appeared on `r round(mean(good_prod$n), 0)` forms, despite only being required to appear on at least 9 forms.
Finally, 28 S-CDI uni-lemmas were not present on the American English CDI:WS (italicized in Fig. 3). <!--, suggesting that this method may help us move beyond an English-centric approach to studying early language. -->

Figure 2 shows the average cross-linguistic difficulty of CDI items by semantic category, for both Swadesh and non-Swadesh items. 
The difficulty of Swadesh items generally tracked with that of non-Swadesh items, although there were cases where one or the other was more or less difficult.


```{r difficulty-by-category, fig.env = "figure", fig.pos = "h", fig.height=3.6, fig.width=3.4, fig.align = "center", fig.cap = "Mean cross-linguistic difficulty of CDI words by semantic category, showing that selection of Swadesh concepts broadly maintained representative difficulty. Bars represent bootstrapped 95\\% confidence intervals."}

xldf <- xldf %>% 
  mutate(SwadeshCDI=ifelse(is.element(uni_lemma, good_prod$uni_lemma), 1, 0))

swad_cat <- xldf %>% filter(SwadeshCDI==1) %>% group_by(category, language) %>%
  summarise(n=n()) %>%
  group_by(category) %>%
  summarise(n = mean(n)) %>% arrange(desc(n))
# sum(swad_cat$n) # 216 words..

swad_lex_cat <- xldf %>% filter(SwadeshCDI==1) %>% group_by(lexical_category, language) %>%
  summarise(n=n()) %>%
  group_by(lexical_category) %>%
  summarise(n = mean(n)) %>% arrange(desc(n))
# 93 nouns, 38 predicates, 28 other, 12 function words
# sum(swad_lex_cat$n) # 171 words..

# remove categories that only exist in 1 or 2 languages:
prod_cat <- xldf %>% 
  filter(!is.na(d), !is.na(category),
         !is.element(category, c("final_particles", "directions", "numbers", "articles", "other",
                                 "descriptive_words (adverbs)", "states", "unknown",
                                 "verb_endings", "verb_modifiers",
                                 "classifiers", "locations_quantities_adverbs"))) %>%
  group_by(language, category, SwadeshCDI) %>%
  summarise(d = mean(d), a1 = mean(a1))

# mean variability of each category, across languages
prod_cat %>% mutate(Items = ifelse(SwadeshCDI==1, "Swadesh", "other")) %>%
  group_by(category, Items) %>%
  tidyboot::tidyboot_mean(d, na.rm=T) %>%
  ggplot(aes(x=reorder(category, mean), y=mean, group=Items, color=Items)) + 
  geom_point(alpha=.7, position = position_dodge2(.2)) +
  geom_linerange(aes(ymin=ci_lower, ymax=ci_upper), alpha=.7, position = position_dodge2(.2)) +
  coord_flip() + theme_classic() + ylab("Mean item difficulty") + xlab("Category") +
  theme(#legend.position="bottom", 
        legend.title=element_blank(),
        legend.position = c(.8, .1),
        legend.margin=margin(c(-1,-1,-1,-1)))
```


```{r fig-mean-swadesh-item-diff, include=F, fig.cap=c("Mean item difficulty of Swadesh CDI items per language. Bars represent bootstrapped 95\\% confidence intervals."), fig.height=3.5, fig.width=3.4}

# summary of info for chosen k
swad_info = swad_agg %>% filter(k==9) %>%
  group_by(sublist) %>%
  summarise(N_avail = mean(N_avail),
            sumscore_cor = mean(sumscore_cor),
            test_info=mean(test_info),
            mean_diff= mean(mean_diff),
            mean_sd = mean(mean_sd),
            sd_sd = mean(sd_sd),
            sd_thresh = mean(sd_thresh),
            swad_mean_diff=mean(swad_mean_diff)) # more cols?

swad_d_m <- xldf %>% filter(is.element(uni_lemma, good_prod$uni_lemma)) %>% # 
  group_by(language) %>% 
  tidyboot::tidyboot_mean(d, na.rm=T) %>%
  arrange(desc(mean))


swad_d_m %>% ggplot(aes(x=reorder(language, -mean), y=mean)) + coord_flip() + 
  geom_point() + theme_classic() +
  geom_linerange(aes(ymin=ci_lower, ymax=ci_upper), alpha=.7) +
  theme(legend.position="bottom", legend.title=element_blank(),
        legend.margin=margin(c(2,5,3,2))) + # legend.margin=margin(c(1,5,5,5)) # top right bottom left
  xlab("Language") + ylab("Mean item dificulty")
```



```{r, swadesh-vs-non-swadesh-diff, echo=F}

tmp <- xldf %>% 
  group_by(SwadeshCDI) %>% 
  filter(!is.na(d)) %>%
  summarise(sd_d = sd(d), d = mean(d), 
            sd_a1 = sd(a1), a1 = mean(a1))
# mean difficulty of Swadesh words: .11; mean difficulty of other words: 0.53
# (mean of all: mean(xldf$d, na.rm=T) .34

swad_discrim = t.test(subset(xldf, SwadeshCDI==1)$a1, subset(xldf, SwadeshCDI==0)$a1) 

swad_diff_ttest = t.test(subset(xldf, SwadeshCDI==1)$d, subset(xldf, SwadeshCDI==0)$d)

gen_test_langs <- low_data_langs$Language
```

Comparing the IRT parameters of the S-CDI uni-lemmas to the rest of the items (across all CDI:WS forms) showed that the discrimination parameter (i.e., slope) of the Swadesh items did not significantly differ from the others, suggesting that the Swadesh items could measure ability as well as non-Swadesh uni-lemmas.
However, S-CDI items were significantly easier than other uni-lemmas (mean S-CDI $d=$ `r round(tmp[2,"d"], 2)`, others' mean $d=$ `r round(tmp[1,"d"], 2)`, `r report_ttest(swad_diff_ttest)`).

To limit the likelihood of a ceiling effect, we proposed extending the S-CDI with 10 additionanl uni-lemmas chosen from the original Swadesh (1971) list, which were selected to fill gaps in the S-CDI, including pronouns (_I, you, we, this, that_), quantifiers (_all, many, not_), and question words (_who, what_).
These uni-lemmas are included on a mean of 21 of the 26 forms, but are of greater difficulty (mean $d=0.84$), and cross-linguistic variability (mean $sd(d)=1.36$).


```{r, echo=F, results='asis', include=F}
good_prod2 <- good_prod %>% left_join(xldf %>% distinct(uni_lemma, category, lexical_category))
#sort(table(good_prod2$category), decr=T)
#sort(table(good_prod2$lexical_category), decr=T)
# nouns, other, predicates, function words
# c(119,55,55,30) / 259

# table(subset(xldf, language=="English (American)")$lexical_category) / 680
# function_words     nouns      other    predicates 
#    0.15          0.46      0.1470588      0.24

tab2 <- good_prod2 %>% group_by(category) %>%
  summarise(n=n()) %>%
  arrange(desc(n))

tab2 <- xtable::xtable(tab2, digits=c(0), 
                       caption = "Number of proposed Swadesh CDI items by semantic category.")
print(tab2, type="latex", comment = F, table.placement = "H", include.rownames=FALSE)
```


## Validating the Swadesh CDI

To validate the S-CDI, we first measured how well simulated raw scores from the S-CDI items correlated with full CDI:WS scores.
This metric approximately reflects how reliable the S-CDI would be as a measure of a child's vocabulary.
On average, for the 26 large CDI:WS datasets, the S-CDI's scores were strongly related to the full CDI:WS scores (mean $r=0.996$; $min=0.989$, $max=0.998$; full table on [OSF](https://osf.io/8swhb/?view_only=6f6ab9818f2a4bb288e05ca9e12f540c)).
We compared these correlations against a baseline that simulated an upper bound for how well the Swadesh CDI could be expected to perform: randomly sample $M$ items from the actual CDI:WS of the target language, where $M$ is the number of Swadesh items that are present on the form.
On average, the 26 CDI:WS forms included $M=176$ of the Swadesh items.

Scores from a random subsample of CDI items tends to perform very well at predicting the overall CDI score, as there are no sampling biases related to item difficulty, or cross-linguistic variability in difficulty or inclusion.
However, note that this is _not_ a viable method to create a new CDI, as in a true CDI construction scenario rather than a simulation, the target CDI would not actually exist! 
Thus, if the S-CDI comes close to performing as well as a random sample from manually curated CDIs, we consider it a success.
Indeed, the random tests had a mean correlation with full CDI scores of $r=0.997$, only $\epsilon=0.001$ higher than the S-CDI.

Next, we measured the total test information yielded by the two baselines (recalling that total test information was one criterion for the construction of the S-CDI).
This metric reflects how well the S-CDI would be able to differentiate the ability of children across different ability levels.
As reported above, the S-CDI yielded a mean total test information of 66085, while the random uni-lemmas baseline yielded a total test information of 66374. 
Although test information was slightly lower for the S-CDI, the values were virtually indistinguishable.

## Testing Generalization of the Swadesh CDI 

We then evaluated the S-CDI's performance in a test of generalization to eight more CDI:WS datasets.
For the eight low-data languages, a comparison of simulated S-CDI scores to full CDI:WS revealed that the S-CDI's raw scores were again strongly related (mean $r=0.990$), with an average of $M=153$ S-CDI items appearing per list.
Table 2 shows the results of this comparison, alongside the upper bound random baseline.
Once again, the S-CDI performed nearly as well as a random sample of the actual CDI (random mean $r=0.994$).
With the 10-item extension, the S-CDI's correlation rose to mean $r=0.993$, demonstrating the value of including items from the more difficult (and variable) categories that were underrepresented on the original list.

```{r generalization-test, echo=F, eval=F}

#xx_eng <- run_comparisons(xldf, languages, good_prod$uni_lemma, 
#                          rand_method="unilemmas_english", metrics=c("sumscore_cor","test_info"))
#xx_wt <- run_comparisons(xldf, languages, good_prod$uni_lemma, 
                         rand_method="unilemmas_weighted", metrics=c("sumscore_cor","test_info"))
xx_it <- run_comparisons(xldf, languages, good_prod$uni_lemma, 
                         rand_method="items", metrics=c("sumscore_cor","test_info"))

# with 10-item extension
xx_it_ext <- run_comparisons(xldf, languages, c(swad10, good_prod$uni_lemma), 
                         rand_method="items", metrics=c("sumscore_cor","test_info"))

sumscore_ttest_it <- run_comparison_ttest(xx_it, "sumscore_cor") # run with paired=F to get means
# .996 vs .997
tinfo_ttest_it <- run_comparison_ttest(xx_it, "test_info") # n.s. difference (report?)
# 66085 vs 66368 n.s.

sumscore_ttest_it_ext <- run_comparison_ttest(xx_it_ext, "sumscore_cor") # run with paired=F to get means
# -.0006, p<.001
tinfo_ttest_it_ext <- run_comparison_ttest(xx_it_ext, "test_info") # n.s. difference (report?)
# n.s. 70874 vs 71242



# just the Swadesh
xx <- run_swadesh_comparisons(xldf, languages, good_prod$uni_lemma)
#summary(xx$`Swadesh r`)
#summary(xx$`Rand r`)

all_unis <- bind_rows(xldf, xldf_lowd)
#gen_xx_eng <- run_comparisons(all_unis, gen_test_langs, good_prod$uni_lemma, 
#                              rand_method="unilemmas_english", metrics=c("sumscore_cor"))

# Swadesh
#run_comparison_ttest(gen_xx_eng, metric="sumscore_cor")


gen_xx_it <- run_comparisons(all_unis, low_data_langs$Language, good_prod$uni_lemma, 
                              rand_method="items", metrics=c("sumscore_cor"))

gen_xx_it_ext <- run_comparisons(all_unis, low_data_langs$Language, c(swad10, good_prod$uni_lemma), 
                              rand_method="items", metrics=c("sumscore_cor"))

gen_sumscore_ttest <- run_comparison_ttest(gen_xx_it, "sumscore_cor") # run with paired=F to get means
# n.s. .987 vs .995

# with 10-item extension
gen_sumscore_ttest_ext <- run_comparison_ttest(gen_xx_it_ext, "sumscore_cor") # run with paired=F to get means
# n.s. .990 vs .995

# note that this uses the set of uni-lemmas on these 8 forms; not the full set from the 26 training langs!
gen_xx <- run_swadesh_comparisons(xldf_lowd, low_data_langs$Language, good_prod$uni_lemma)
#summary(gen_xx$`Swadesh r`) 
# Min.       Mean     Max. 
# 0.9825    0.990   0.997 

#summary(gen_xx$`Rand r`)
#  Min.       Mean      Max. 
# 0.987     0.995    0.997 

### Add some difficult Swadesh words, from underrepresented S-CDI categories
# (pronouns, question words)
#"I","you","we","this","that","who","what","not","all","many" 
swad10 <- c("1SG","2SG","1PL","this","that","who","what","not","all","many")

#low_var <- uni_per_form %>% filter(!is.element(uni_lemma, good_prod$uni_lemma), d_sd<1.18, n>=9)

View(subset(uni_per_form, is.element(uni_lemma, swad10)))

xx10 <- run_swadesh_comparisons(xldf, languages, c(swad10, good_prod$uni_lemma))
mean(xx10$`Swadesh r`)
# .996 # 190 available vs 176 

gen_xx10 <- run_swadesh_comparisons(xldf_lowd, low_data_langs$Language, c(swad10, good_prod$uni_lemma))
mean(gen_xx10$`Swadesh r`)
# .993 vs 

save(xx, xx_it, #xx_eng, xx_wt,
     xx10, xx_it_ext,
     gen_xx, gen_xx10, # gen_xx_eng, gen_xx_wt,
     file=here("data/Swadesh_comparisons.Rdata"))
```

```{r gen-results, results='asis'}
load(here("data/Swadesh_comparisons.Rdata"))
# mean(gen_xx10$`Swadesh r`) # .993

gen_xx <- gen_xx %>% rename(`S-CDI r` = `Swadesh r`) 
gen_xx$`S-CDI+10 r` = gen_xx10$`Swadesh r` # get the Swad + 10 extension r

gen_tab <- xtable::xtable(gen_xx, digits=c(3), 
                       caption = "Generalization test results for S-CDI vs. random baseline, and extended S-CDI (+10 items).")
print(gen_tab, type="latex", comment = F, table.placement = "H",
      include.rownames=FALSE, size="\\fontsize{9pt}{10pt}\\selectfont")
```



```{r, include=F}
short_wsA <- read_csv(here("data/eng-ws-shortA.csv"))
short_wsB <- read_csv(here("data/eng-ws-shortB.csv"))

wsA_int = sort(intersect(short_wsA$word, good_prod$uni_lemma)) # 28
# "airplane" "all gone" "bed" "beside"   "broom"    "candy"    "cat" "chin" "cold" "dog" "duck"  
# "fit" "hand"     "horse"    "like"     "meow"     "milk"     "necklace" "ouch" "peas" "school" "shopping"
# "sky" "stairs"   "star"     "tonight"  "towel"    "under"  
wsB_int = sort(intersect(short_wsB$word, good_prod$uni_lemma)) # 26
#  "before" "bite" "black" "careful" "catch" "cloud" "dirty"  
# "drink (action)" "duck" "hose" "jump" "kick" "mailman" "nose"          
# "ouch" "penguin" "quack quack" "salt" "sun" "telephone" "then"          
# "today" "tomorrow" "tongue" "yum yum" "zoo" 

# intersect(wsA_int, wsB_int) # 2 - "duck" "ouch"

#sort(short_wsA$word)
#sort(good_prod$uni_lemma)
# maybe remove? penis, vagina, on the basis of DIF
```



# Discussion

This study compared psychometric models fitted to 26 CDI datasets in order to find concepts that had low variability in their cross-linguistic difficulty, and that were frequently included on CDI:WS forms. 
We identified 229 concepts that appeared on at least 9 of the CDIs, and which had more consistent cross-linguistic difficulty than other concepts appearing on multiple CDIs.
Using real-data simulations, we showed that administering this set of Swadesh CDI items would generate scores that were strongly related to full CDI:WS scores, both for the original 26 datasets, and in a generalization test to eight low-data languages.
Moreover, the Swadesh CDI items resulted in comparable total test information to tests of the same length composed of randomly-selected uni-lemmas from the target test -- a challenging baseline to beat, and a construction method impossible to use when creating a new CDI.
The Swadesh CDI contains items with relatively stable cross-linguistic difficulty estimates, and in the absence of access to researchers who are familiar with relevant local cultures and concepts, they may serve as a rapid, simple means of approximating children's ability levels, even in the absence of a large norming dataset.

However, the Swadesh CDI items were also significantly easier than other items, meaning that older children may perform at ceiling if given only the Swadesh CDI items.
(This may be unsurprising from the perspective that Swadesh words are meant to be universal, and are therefore more frequent and basic---both within and across individual children's experiences.)
Thus, our suggested use case for the Swadesh CDI list is as a starting point for researchers seeking to develop a CDI in a new language, rather than as a complete short-form CDI based on existing long-form CDI data.
In particular, researchers should seek to add relevant uni-lemmas from categories that were less well-represented on the S-CDI, including question words, quantifiers, helping verbs, and pronouns. 
These categories also tended to be more difficult, so adding items from them is likely to increase the difficulty ceiling of the form. 
Indeed, the inclusion of 10 such items drawn from the original Swadesh (1971) list increased generalization performance.

Another potential limitation of this work is that most existing CDIs (and most datsets available in Wordbank) target languages in the Indo-European language family. <!-- add sentence about cultural bias -->
It is not clear to what extent this bias in the existing data might interfere with generalizing to non-Indo-European languages.
Nonetheless, our original 26 datasets include 7 non-Indo-European languages (3 Sino-Tibetan, 1 Afro-Asiatic, 1 Uralic, 1 Koreanic, 1 Turkic), and the generalization datasets include 1 Uralic and 2 Niger-Congo languages (a language family not represented in the original datasets); the broad consistency across language families thus suggests that the effectiveness of the S-CDI may be sufficiently robust.
<!--Such a robustness also corroborates with results from @floccia2018, who demonstrated that a bilingual vocabulary model based on a set of 30 uni-lemmas had good predictive validity on non-target languages, even those from other language families.-->

Developing a list of appropriate vocabulary words is not the only challenge researchers face when seeking to develop and use parent-report measures in a new language and culture. 
The pragmatics of language between children and adults can differ greatly across cultures, and has been found to interfere with administration of parent-report measures of early vocabulary, for example in Kiswahili [@alcock2017production] and Wolof [@weber2018]. <!-- TODO: FOR EXAMPLE... -->
As such, local cultural knowledge remains essential in appropriately developing and administering novel CDI adaptations.

Despite the myriad challenges that remain in creating new measures of early language development, we believe that the proposed Swadesh CDI list will give researchers a solid foundation to start from, lowering the barrier to the adaptation of CDI forms in new languages, since these are often time-consuming and challenging to construct.
Expanding the number of languages with effective vocabulary measures would be a critical step in addressing issues related to the under-representation of lingustic diversity in language acquisition research [@kidd2022]. 
Certainly, increasing the diversity of languages studied is a critical step towards developing a truly general understanding of how young children learn language.

# Acknowledgements

Redacted for anonymous review.
<!-- We would like to thank all of the contributors to Wordbank, from the researchers who created and adapted the CDIs to those who collected the data (as well as the participants), to those who have created and maintained Wordbank over the years. -->

# References 

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent
